---
title: "Betsy Rosalen HW 2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HW 2

________________________________________________________________________________

## Question 1

There are 540 identical plastic chips numbered 1 through 540 in a box.  What is the probability of reaching into the box and randomly drawing the chip numbered 505?  Express your answer as a fraction or a decimal number rounded to four decimal places.

$$P(505) = \frac{1}{540}$$
 
________________________________________________________________________________

## Question 2

Write out the sample space for the given experiment.  Separate your answers using commas.

When deciding what you want to put into a salad for dinner at a restaurant, you will choose one of the following extra toppings: asparagus, cheese.  Also, you will add one of following meats: eggs, turkey.  Lastly, you will decide on one of the following dressings: French, vinaigrette.  (Note: Use the following letters to indicate each choice: A for asparagus, C for cheese, E for eggs, T for turkey, F for French, and V for vinaigrette.)

$$\text{S = {AEF, AEV, ATF, ATV, CEF, CEV, CTF, CTV}}$$
 
________________________________________________________________________________

## Question 3

A card is drawn from a standard deck of 52 playing cards.  What is the probability that the card will be a heart and not a face card?  Write your answer as a fraction or a decimal number rounded to four decimal places.
 $$P(\text{heart and not a face card}) = \frac{10}{52} \text{ or reduced to } \frac{5}{26}$$ 
 
________________________________________________________________________________

## Question 4

A standard pair of six-sided dice is rolled.  What is the probability of rolling a sum less than 6?  Write your answer as a fraction or a decimal number rounded to four decimal places.
$$P(sum < 6) = \frac{10}{36} \text{ or } \frac{5}{18}$$
 
________________________________________________________________________________

## Question 5

A pizza delivery company classifies its customers by gender and location of residence.  The research department has gathered data from a random sample of 2001 customers.  The data is summarized in the table below.
	
Gender and Residence of Customers | | 
----- | ----- | -----
---	| Males	| Females
Apartment	| 233	| 208
Dorm	| 159	| 138
With Parent(s)	| 102	| 280
Sorority/Fraternity House	| 220	| 265
Other	| 250	| 146

What is the probability that a customer is male?  Write your answer as a fraction or a decimal number rounded to four decimal places. 

$$P(\text{Male}) = \frac{233+159+102+220+250}{2001}$$
$$P(\text{Male}) = \frac{964}{2001}$$

________________________________________________________________________________

## Question 6

Three cards are drawn with replacement from a standard deck.   What is the probability that the first card will be a club, the second card will be a black card, and the third card will be a face card?  Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(\text{Club})\cdot P(\text{Black})\cdot P(\text{Face}) = \frac{4}{52}\cdot \frac{26}{52}\cdot \frac{12}{52} = \frac{3}{338}$$
 
________________________________________________________________________________

## Question 7

Two cards are drawn without replacement from a standard deck of 52 playing cards. 

What is the probability of choosing a spade for the second card drawn, if the first card, drawn without replacement, was a heart?  Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(\text{Spade | Heart}) = \frac{13}{51}$$
 
________________________________________________________________________________

## Question 8

Two cards are drawn without replacement from a standard deck of 52 playing cards. 

What is the probability of choosing a heart and then, without replacement, a red card?  Write your answer as a fraction or a decimal number rounded to four decimal places.

$$P(\text{Heart})\cdot P(\text{Red | Heart}) = \frac{13}{52} \cdot \frac{25}{51} = \frac{25}{204}$$
 
________________________________________________________________________________

## Question 9

There are 85 students in a basic math class.  The instructor must choose two students at random.
	
Students in a Basic Math Class |  |  
----- | ----- | -----
--- | Males | Females
Freshmen	 | 12	 | 12
Sophomores	 | 19	 | 15
Juniors	 | 12	 | 4
Seniors	 | 7	 | 4


What is the probability that a junior female and then a freshmen male are chosen at random?  Write your answer as a fraction or a decimal number rounded to four decimal places. 

$$P(\text{JF})\cdot P(\text{FM | JF}) = \frac{4}{85} \cdot \frac{12}{84} = \frac{4}{595}$$

________________________________________________________________________________

## Question 10

Out of 300 applicants for a job, 141 are male and 52 are male and have a graduate degree.

**Step 1.** What is the probability that a randomly chosen applicant has a graduate degree, given that they are male?  Enter your answer as a fraction or a decimal rounded to four decimal places. 

$$P(\text{Graduate Degree | Male}) = \frac{52}{141}$$

**Step 2.** If 102 of the applicants have graduate degrees, what is the probability that a randomly chosen applicant is male, given that the applicant has a graduate degree?  Enter your answer as a fraction or a decimal rounded to four decimal places. 

$$P(\text{Male | Graduate Degree}) = \frac{52}{102}$$

________________________________________________________________________________

## Question 11

A value meal package at Ron's Subs consists of a drink, a sandwich, and a bag of chips.  There are 6 types of drinks to choose from, 5 types of sandwiches, and 3 types of chips.  How many different value meal packages are possible?

$$ 6 \cdot 5 \cdot 3 = \text{ 90 different value meal packages are possible.}$$

________________________________________________________________________________

## Question 12

A doctor visits her patients during morning rounds.  In how many ways can the doctor visit 5 patients during the morning rounds?

```{r}
factorial(5)
```
The doctor can visit her patients in 120 different ways during her morning rounds.

________________________________________________________________________________

## Question 13

A coordinator will select 5 songs from a list of 8 songs to compose an event's musical entertainment lineup.  How many different lineups are possible?

```{r}
fp=function(N,n){factorial(N)/factorial(N-n)}
fp(8,5)
```

6720 different lineups are possible assuming that the order of songs matters.

________________________________________________________________________________

## Question 14

A person rolls a standard six-sided die 9 times.  In how many ways can he get 3 fours, 5 sixes and 1 two?

```{r}
fc=function(N,n){factorial(N)/(factorial(n)*factorial(N-n))}
fours <- fc(9,3)
fours
sixes <- fc(6,5)
sixes
fours * sixes
```

There are 504 ways to get 3 fours, 5 sixes and 1 two if you roll a die 9 times.

________________________________________________________________________________

## Question 15

How many ways can Rudy choose 6 pizza toppings from a menu of 14 toppings if each topping can only be chosen once?

```{r}
fc(14,6)
```

There are 3003 different combinations of 6 toppings that Rudy can order for his pizza.

________________________________________________________________________________

## Question 16

3 cards are drawn from a standard deck of 52 playing cards.  How many different 3-card hands are possible if the drawing is done without replacement?

```{r}
fc(52,3)
```

There are 22,100 different 3 card hands possible from a 52 card deck.

________________________________________________________________________________

## Question 17

You are ordering a new home theater system that consists of a TV, surround sound system, and DVD player.  You can choose from 12 different TVs, 9 types of surround sound systems, and 5 types of DVD players.  How many different home theater systems can you build?

$$ 12 \cdot 9 \cdot 5 = \text{ 540 different home theater systems are possible.}$$

________________________________________________________________________________

## Question 18

You need to have a password with 5 letters followed by 3 odd digits between 0 - 9 inclusively.  If the characters and digits cannot be used more than once, how many choices do you have for your password?

```{r}
A = fp(52,5)  # Assuming case sensitive so there are 26 lowercase and 26 uppercase characters possible
B = fp(10,3)
A*B

a = fp(26,5)  # Assuming NOT case sensitive so there are 26 letters
b = fp(10,3)
a*b
```

There are 224,550,144,000 possible passswords assuming that the password is case sensitive and 5,683,392,000 if it is not case sensitive.

________________________________________________________________________________

## Question 19

Evaluate the following expression.
	 $$_9 P_4$$

```{r}
fp(9,4)
```


________________________________________________________________________________

## Question 20

Evaluate the following expression.
	 $$_{11} C_8$$

```{r}
fc(11,8)
```


________________________________________________________________________________

## Question 21

Evaluate the following expression.
	$$\frac{ _{12} P_8}{_{12} C_4}$$

```{r}
N = fp(12,8)
D = fc(12,4)
N/D
```


________________________________________________________________________________

## Question 22

The newly elected president needs to decide the remaining 7 spots available in the cabinet he/she is appointing.  If there are 13 eligible candidates for these positions (where rank matters), how many different ways can the members of the cabinet be appointed?

```{r}
fp(13,7)
```

There are 8,648,640 different ways the members of the cabinet can be appointed.

________________________________________________________________________________

## Question 23

In how many ways can the letters in the word 'Population' be arranged?

Assuming the uppercase P and lowercase p are the same...

```{r}
p <- fc(10,2)
p
o <- fc(8,2)
o
rest <-fp(6,6)
rest
p*o*rest
```

There are 907,200 ways that you can arrange the letters in the word population if your word is in all upper or lower case or not case-sensitive

If the letters ARE case-sensitive and the two P's are different...

```{r}
o <- fc(10,2)
o
rest <-fp(8,8)
rest
o*rest
```

There are 1,814,400 ways to arrange the letters if you treat the upper and lower case P's as two different letters.

________________________________________________________________________________

## Question 24

Consider the following data:
	
**x** |	5 |	6 |	7 |	8 |	9
----- |	----- |	----- |	----- |	----- |	-----
**p(x)** |	0.1 |	0.2 |	0.3 |	0.2 |	0.2


**Step 1.** Find the expected value $E(X)$.  Round your answer to one decimal place.  

```{r}
mu <- 5*0.1 + 6*0.2 + 7*0.3 + 8*0.2 + 9*0.2
mu
```

$$E(X) = 7.2$$

**Step 2.** Find the variance.  Round your answer to one decimal place. 

```{r}
v = round(((5-mu)^2)*0.1 + ((6-mu)^2)*0.2 + ((7-mu)^2)*0.3 + ((8-mu)^2)*0.2 + ((9-mu)^2)*0.2,1)
v
```

$$V(X) = 1.6$$

**Step 3.** Find the standard deviation.  Round your answer to one decimal place. 

```{r}
sd <- round(sqrt(v),1)
sd
```

$$\sigma = 1.3$$

**Step 4.** Find the value of $P(X ³ 9)$.  Round your answer to one decimal place.  


**Step 5.** Find the value of $P(X £ 7)$.  Round your answer to one decimal place.  

________________________________________________________________________________

## Question 25

Suppose a basketball player has made 188 out of 376 free throws.  If the player makes the next 3 free throws, I will pay you \$23.  Otherwise you pay me \$4.

**Step 1.** Find the expected value of the proposition.  Round your answer to two decimal places. 

```{r}
x = 188/376
x
P = x^3
P
EV = P*23 + (1-P)*-4
round(EV, 2)
```
$$\text{Expected Value} = -\$ 0.62$$

**Step 2.** If you played this game 994 times how much would you expect to win or lose? (Losses must be entered as negative.) 

```{r}
994 * EV
```

I'd expect to lose -\$621.25 if I played the game 994 times.

________________________________________________________________________________

## Question 26

Flip a coin 11 times.  If you get 8 tails or less, I will pay you \$1.  Otherwise you pay me \$7.

**Step 1.** Find the expected value of the proposition.  Round your answer to two decimal places. 

```{r}
P = pbinom(8,11,.5)
P
EV = round(P*1 + (1-P)*-7, 2)
EV
```

$$\text{Expected Value} = \$ 0.74$$ 

**Step 2.** If you played this game 615 times how much would you expect to win or lose? (Losses must be entered as negative.) 

```{r}
615 * EV
```

I'd expect to win \$455.10 if I played the game 615 times.

________________________________________________________________________________

## Question 27

If you draw two clubs on two consecutive draws from a standard deck of cards you win \$583.  Otherwise you pay me \$35.  (Cards drawn without replacement.)

**Step 1.** Find the expected value of the proposition.  Round your answer to two decimal places. 

```{r}
P = 13/52 * 12/51
P
EV = round(P*583 + (1-P)*-35, 2)
EV
```
$$\text{Expected Value} = \$ 1.35$$

**Step 2.** If you played this game 632 times how much would you expect to win or lose? (Losses must be entered as negative.) 

```{r}
632 * EV
```

I'd expect to win \$853.20 if I played the game 632 times.

________________________________________________________________________________

## Question 28

A quality control inspector has drawn a sample of 10 light bulbs from a recent production lot.  If the number of defective bulbs is 2 or less, the lot passes inspection.  Suppose 30% of the bulbs in the lot are defective.  What is the probability that the lot will pass inspection? (Round your answer to 3 decimal places)

Assuming that the sample is less than 1/20th of the production lot

```{r}
round(ppois(2,3),3)
```


________________________________________________________________________________

## Question 29

A quality control inspector has drawn a sample of 5 light bulbs from a recent production lot.  Suppose that 30%  of the bulbs in the lot are defective.  What is the expected value of the number of defective bulbs in the sample?  Do not round your answer. 

1.5 out of a sample of 5 light bulbs would be expected to be deffective.

________________________________________________________________________________

## Question 30

The auto parts department of an automotive dealership sends out a mean of 5.5 special orders daily. What is the probability that, for any day, the number of special orders sent out will be more than 5? (Round your answer to 4 decimal places)

```{r}
round(1 - ppois(5,5.5),4)
```

________________________________________________________________________________

## Question 31

At the Fidelity Credit Union, a mean of  5.7 customers arrive hourly at the drive-through window.  What is the probability that, in any hour, more than 4 customers will arrive? (Round your answer to 4 decimal places)

```{r}
round(1 - ppois(4,5.7),4)
```

________________________________________________________________________________

## Question 32

The computer that controls a bank's automatic teller machine crashes a mean of  0.4 times per day. What is the probability that, in any 7-day week, the computer will crash no more than 1 time? (Round your answer to 4 decimal places)

```{r}
.4*7
round(ppois(1,2.8),4)
```

________________________________________________________________________________

## Question 33

A town recently dismissed 8 employees in order to meet their new budget reductions.  The town had 6 employees over 50 years of age and 19 under 50.  If the dismissed employees were selected at random, what is the probability that more than 1 employee was over 50?  Write your answer as a fraction or a decimal number rounded to three decimal places.

```{r}
M = 6/25
P = round(1 - pbinom(1,8,M),3)
P
```

There is a 0.608 probability that more than one employee was over 50.

________________________________________________________________________________

## Question 34

Unknown to a medical researcher, 10 out of 25 patients have a heart problem that will result in death if they receive the test drug.  Eight patients are randomly selected to receive the drug and the rest receive a placebo.  What is the probability that less than 7 patients will die?  Write your answer as a fraction or a decimal number rounded to three decimal places.

```{r}
M = 10/25
P = round(pbinom(6,8,M),3)
P
```

Wow!  Doesn't seem very ethical!